---
title: "Draft raster methodology in R"
author:
- Jeff Hollister
- Q Kellogg
date: '2017-11-27'
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r}
devtools::install_github("jimhester/archive")
library(here)
library(archive)
library(httr)
library(raster)
library(sp)
library(rgdal)
library(tidyverse)
library(sf)
```
This is a working script to be used to figure out the raster methodology for
N-Sink.  It will be messy.

# Step 0: Get Data
## Data
Grab flow direction and 12 Digist HUSCS from NHD Plus

```{r}
# Grab data if doesn't exist locally
if(!file.exists(here("data/nhdplus_ne_fdr.7z"))){
  fdr_url <- "http://www.horizon-systems.com/NHDPlusData/NHDPlusV21/Data/NHDPlusNE/NHDPlusV21_NE_01_01a_FdrFac_01.7z"
  httr::GET(fdr_url, httr::write_disk(path = here("data/nhdplus_ne_fdr.7z")),progress())
  archive_extract(here("data/nhdplus_ne_fdr.7z"), here("data"))
}
if(!file.exists(here("data/nhdplus_ne_huc12.7z"))){
  huc12_url <- "http://www.horizon-systems.com/NHDPlusData/NHDPlusV21/Data/NHDPlusNE/NHDPlusV21_NE_01_WBDSnapshot_03.7z"
  httr::GET(huc12_url, httr::write_disk(here("data/nhdplus_ne_huc12.7z")), httr::progress())
  archive_extract(here("data/nhdplus_ne_huc12.7z"), here("data"))
}
fdr_ne <- raster(here("data/NHDPlusNE/NHDPlus01/NHDPlusFdrFac01a/fdr/"))
huc12_ne <- st_read(here("data/NHDPlusNE/NHDPlus01/WBDSnapshot/WBD/WBD_Subwatershed.shp")) %>% st_transform(projection(fdr_ne))
niantic_huc <- huc12_ne %>%
  filter(HU_12_NAME == "Niantic River") 
niantic_fdr <- crop(fdr_ne, as(niantic_huc,"Spatial"))
```


# Step 1: Generate flow path from user specified location


NHD Plus Flow Direction
TauDEM?
raster::flowPath()

```{r}
plot(as(niantic_huc, "Spatial"))
loc <- unlist(locator(1))
example_fp <- flowPath(niantic_fdr, loc)
example_fp_xy <- xyFromCell(niantic_fdr, example_fp)
plot(niantic_fdr)
plot(niantic_huc, add = T)
lines(example_fp_xy)
```

# Step 2: Summarize land, stream, and lake removal 

land removal is a constant made up of pixels that are vegetated Hydric Soils
Combination of hydric soils from SSURGO and Vegetated land cover from NLCD
Figuring this layer(s) out will be the most work.

## Data
https://github.com/ncss-tech and soilDB

## Output

What does the output for this look like? Table? Flow path with removal values?  Something else?

Bare minimum is table of removal.  Per segment ideally.

Example output #1 - landcover summary table - Along a 100 m buffer

| Landcover Class | Total Area | 
|-----------------|------------|
| Developed       | XX         |
| Forested        | XX         |
| Wetland         | XX         |
| Other           | XX         |

Example output #2 - Nitrogen removal per segment (drop to integer values)

| Segement Type | Length | Percent Removal | N in | N out |
|---------------|--------|-----------------|------|-------|
| Non-hydric    | 8156   | 0               | 100  | 100   |
| Hydric        | 546    | 80              | 100  | 20    |
| Stream        | XX     | 16              | 20   | 16.8  |
| Lake/Pond     | XX     | 23              | 16.8 | 12.9  |
| Stream        | XX     | 1               | 12.9 | 11.6  |


